### 2018-03-20 Sections/quantum_algorithms.tex
Phase est alg complete

\begin{table}[!htb]
\fbox{\parbox{0.965\columnwidth}{\texttt{ 
function PhaseEstimation($\hat{U}$):
\begin{enumerate}
	\item $\Box$
\end{enumerate}
\begin{align}
\Qcircuit @C=1em @R=1.6em {
    \lstick{\ket{0}} & \gate{\hat{H}} & \qw &\qw & \dots && \ctrl{3} & \multigate{2}{\hat{QFT}^{-1}} & \meter \\
    \lstick{\vdots} & \gate{\hat{H}} & \qw &\qw & \dots & & \qw & \ghost{\hat{QFT}^{-1}} & \meter \\
    \lstick{\ket{0}} & \gate{\hat{H}} & \ctrl{1} &\qw & \dots && \qw & \ghost{\hat{QFT}^{-1}} & \meter \\
    \lstick{\ket{\psi}} & \qw & \gate{\hat{U}^{2^0}} &\qw & \dots & &\gate{\hat{U}^{2^{n-1}}} & \qw & \qw \\
} \nonumber
\end{align}
}}}
\caption{The quantum phase-estimation algorithm.} \label{alg:q_phase_est}
\end{table}

@@

\subsection{Integer factorisation} \index{Shor's algorithm}\index{Hidden subgroup problem}\index{Period-finding algorithm}\index{Discrete logarithm algorithm}

@@

\subsection{Topological data analysis} \index{Topological data analysis}

>>>>

The goal of the phase-estimation algorithm \cite{bib:PhaseEstAlg} is to calculate the eigenvalue of a given unitary operator for a given eigenvector of that unitary. Since the eigenvalues of unitary operators are always phases of amplitude 1, we can write,
\begin{align}
	\hat{U}\ket\psi = e^{2\pi i\theta}\ket\psi,
\end{align}
where $\theta$ parameterises the phase of the eigenvalue and $\ket\psi$ is an eigenvector of $\hat{U}$.

The algorithm for this is baked from several ingredients: a Hadamard transform\index{Hadamard transform}; controlled $\hat{U}$s\footnote{Any unitary has a coherently controlled unitary equivalent, of the form $\hat{U}_\text{controlled} = \ket{0}\bra{0}\otimes \hat{\mathbb{I}} + \ket{1}\bra{1}\otimes\hat{U}$. These always have an efficient circuit construction, given $\hat{U}$.}; and, an inverse quantum Fourier transform (QFT)\index{Quantum Fourier transform}.

The QFT matrix contains elements taken from the discrete Fourier transform (DFT)\index{Discrete Fourier transform}. Specifically, the \mbox{$N\times N$} QFT matrix has elements,
\begin{align}
\hat{QFT}_{j,k} = \frac{1}{\sqrt{N}} \omega^{(j-1)(k-1)},	
\end{align}
where,
\begin{align}
\omega = e^{\frac{2\pi i}{N}},	
\end{align}
is a complex root of unity. Equivalently, this implements the map,
\begin{align}
	\hat{QFT} \ket{j} \to \frac{1}{\sqrt{N}}\sum_{k=0}^{N-1} \omega^{jk}\ket{k},
\end{align}
on the $N$ basis states. The QFT of any dimension has an efficient circuit implementation, making it an important subroutine in many algorithms.

Alg.~\ref{alg:q_phase_est} shows the circuit implementation of the phase-estimation algorithm. The output is a binary representation of \mbox{$2^n\theta$}, where the precision is determined entirely by the number of qubits in the output register, which can be arbitrarily large in principle.

\begin{table}[!htb]
\fbox{\parbox{0.965\columnwidth}{\texttt{ 
function PhaseEstimation($\hat{U}$, $\ket\psi$):
\begin{enumerate}
	\item For unitary $\hat{U}$ with eigenvector $\ket\psi$, apply the circuit below.
	\item The register of qubits at the output to the QFT encodes $2^n\theta$ in binary representation.
	\item $\Box$
\end{enumerate}
\begin{align}
\Qcircuit @C=1em @R=1.6em {
    \lstick{\ket{0}} & \gate{\hat{H}} & \qw &\qw & \dots && \ctrl{3} & \multigate{2}{\hat{QFT}^{-1}} & \meter \\
    \lstick{\vdots} & \gate{\hat{H}} & \qw &\qw & \dots & & \qw & \ghost{\hat{QFT}^{-1}} & \meter \\
    \lstick{\ket{0}} & \gate{\hat{H}} & \ctrl{1} &\qw & \dots && \qw & \ghost{\hat{QFT}^{-1}} & \meter \\
    \lstick{\ket{\psi}} & \qw & \gate{\hat{U}^{2^0}} &\qw & \dots & &\gate{\hat{U}^{2^{n-1}}} & \qw & \qw \\
} \nonumber
\end{align}
}}}
\caption{The quantum phase-estimation algorithm.} \label{alg:q_phase_est}
\end{table}

The phase-estimation algorithm is not extremely useful as a standalone product, but acts as a necessary subroutine in many important algorithms, such as Shor's algorithm\index{Shor's algorithm} (Sec.~\ref{sec:shors_alg}) and topological data analysis\index{Topological data analysis} (Sec.~\ref{sec:TDA}).

@@

\subsection{Integer factorisation} \index{Shor's algorithm}\index{Hidden subgroup problem}\index{Period-finding algorithm}\index{Discrete logarithm algorithm}\label{sec:shors_alg}

@@

\subsection{Topological data analysis} \index{Topological data analysis}\label{sec:TDA}
